---
title: "Stats_Summative"
author: "Briggs"
date: "11/16/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Import Data

```{r}
library(readxl)

HighGHI <- read_excel("C:/Users/mtbri/OneDrive - Nexus365/Summatives/HighGHI_12_27.xlsx")
MidGHI <- read_excel("C:/Users/mtbri/OneDrive - Nexus365/Summatives/Mid_GHI_12_27.xlsx")
```


## Summary of variables in HIGH data set


```{r}
summary(HighGHI)
table(is.na(HighGHI))
```

## Summary of variables in MID data set


```{r}
summary(MidGHI)
table(is.na(MidGHI))
```

## Dropping NAs and making variables numeric 


```{r}
HighGHI_new = HighGHI
HighGHI_new$Country <- NULL


HighGHI_new <- na.omit(HighGHI_new) 
table(is.na(HighGHI_new))

summary(HighGHI_new)

HighGHI_new$AVG_Dietary_Esupply <- as.numeric(HighGHI_new$AVG_Dietary_Esupply)

HighGHI_new$AVG_Food_Production <- as.numeric(HighGHI_new$AVG_Food_Production)

HighGHI_new$AVG_Protein_Supply <- as.numeric(HighGHI_new$AVG_Protein_Supply)

HighGHI_new$AVG_Protein_Supply_AO <- as.numeric(HighGHI_new$AVG_Protein_Supply_AO)

HighGHI_new$GDP <- as.numeric(HighGHI_new$GDP)

HighGHI_new$Undernourishment <- as.numeric(HighGHI_new$Undernourishment)

HighGHI_new$Cereal_Import_Dependency <- as.numeric(HighGHI_new$Cereal_Import_Dependency)

HighGHI_new$PCT_Arable_Land <- as.numeric(HighGHI_new$PCT_Arable_Land)

HighGHI_new$Imports_vs_Exports <- as.numeric(HighGHI_new$Imports_vs_Exports)

HighGHI_new$Political_Stability <- as.numeric(HighGHI_new$Political_Stability)

HighGHI_new$PC_Food_Production_Variability <- as.numeric(HighGHI_new$PC_Food_Production_Variability)


HighGHI_new$PC_Food_Supply_Variability<- as.numeric(HighGHI_new$PC_Food_Supply_Variability)

HighGHI_new$PCT_Basic_Drinking_Water <- as.numeric(HighGHI_new$PCT_Basic_Drinking_Water)

HighGHI_new$PCT_Obesity_18Older <- as.numeric(HighGHI_new$PCT_Obesity_18Older)

HighGHI_new$PCT_Anemia_Women <- as.numeric(HighGHI_new$PCT_Anemia_Women)

```

## Dropping NAs

```{r}

MidGHI_new = MidGHI


MidGHI_new <- na.omit(MidGHI_new) 
table(is.na(MidGHI_new))

summary(MidGHI_new)



```


## Making data numeric for moderate to low GHI levels

```{r}
MidGHI_new$AVG_Dietary_Esupply <- as.numeric(MidGHI_new$AVG_Dietary_Esupply)

MidGHI_new$AVG_Food_Production <- as.numeric(MidGHI_new$AVG_Food_Production)

MidGHI_new$AVG_Protein_Supply <- as.numeric(MidGHI_new$AVG_Protein_Supply)

MidGHI_new$AVG_Protein_Supply_AO <- as.numeric(MidGHI_new$AVG_Protein_Supply_AO)

MidGHI_new$GDP <- as.numeric(MidGHI_new$GDP)

MidGHI_new$Undernourishment <- as.numeric(MidGHI_new$Undernourishment)

MidGHI_new$Cereal_Import_Dependency <- as.numeric(MidGHI_new$Cereal_Import_Dependency)

MidGHI_new$PCT_Arable_Land <- as.numeric(MidGHI_new$PCT_Arable_Land)

MidGHI_new$Imports_vs_Exports <- as.numeric(MidGHI_new$Imports_vs_Exports)

MidGHI_new$Political_Stability <- as.numeric(MidGHI_new$Political_Stability)

MidGHI_new$PC_Food_Production_Variability <- as.numeric(MidGHI_new$PC_Food_Production_Variability)


MidGHI_new$PC_Food_Supply_Variability<- as.numeric(MidGHI_new$PC_Food_Supply_Variability)

MidGHI_new$PCT_Basic_Drinking_Water <- as.numeric(MidGHI_new$PCT_Basic_Drinking_Water)

MidGHI_new$PCT_Obesity_18Older <- as.numeric(MidGHI_new$PCT_Obesity_18Older)

MidGHI_new$PCT_Anemia_Women <- as.numeric(MidGHI_new$PCT_Anemia_Women)
```


## Plots for Exploratory Data Analysis


```{r}
library(ggplot2)
library(ggthemes)
library(cowplot)



p1 <- ggplot(HighGHI_new, aes(x = AVG_Dietary_Esupply, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Average Dietary Energy Supply") + geom_smooth(method=lm, col= "blue") 

#png(height=400, width=800, filename="ggplots1", type="cairo")


p2 <- ggplot(HighGHI_new, aes(x = AVG_Food_Production, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Average Food Production") + geom_smooth(method=lm, col="blue")

p3 <- ggplot(HighGHI_new, aes(x = PC_Food_Production_Variability, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Percentage of Food Production Variability") + geom_smooth(method=lm, col="blue")



p4 <- ggplot(HighGHI_new, aes(x = AVG_Protein_Supply, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Average Protein Supply") + geom_smooth(method=lm, col="blue")


p5 <- ggplot(HighGHI_new, aes(x = AVG_Protein_Supply_AO, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Average Protein Supply of Animal Origin") + geom_smooth(method=lm, col="blue")


p6 <- ggplot(HighGHI_new, aes(x = PC_Food_Supply_Variability, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Percentage of Food Supply Variability") + geom_smooth(method=lm, col="blue")

p7 <- ggplot(HighGHI_new, aes(x = GDP, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "GDP") + geom_smooth(method=lm, col="blue")


p8 <- ggplot(HighGHI_new, aes(x = Cereal_Import_Dependency, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Cereal Import Dependency") + geom_smooth(method=lm, col="blue")

p9 <- ggplot(HighGHI_new, aes(x = Undernourishment, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Undernourishment") + geom_smooth(method=lm, col="blue")


p10 <- ggplot(HighGHI_new, aes(x = PCT_Arable_Land, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Percentage of Arable Land") + geom_smooth(method=lm, col="blue")

p11 <- ggplot(HighGHI_new, aes(x = Imports_vs_Exports, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Imports vs. Exports") + geom_smooth(method=lm, col="blue")


p12 <- ggplot(HighGHI_new, aes(x = PCT_Basic_Drinking_Water, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Percentage of Basic Drinking Water Available") + geom_smooth(method=lm, col="blue")

p13 <- ggplot(HighGHI_new, aes(x = PCT_Obesity_18Older, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Percentage of Obesity in 18 and Older Individuals") + geom_smooth(method=lm, col="blue")

p14 <- ggplot(HighGHI_new, aes(x = PCT_Anemia_Women, y = Political_Stability)) +
    geom_point(col = "black") + labs(y= "Political Stability", x= "Percentage of Women with Anemia") + geom_smooth(method=lm, col="blue")





plot_grid(p1,p2,p3,p4)

```



```{r}
#png(height=400, width=800, filename="ggplots2", type="cairo")
plot_grid(p5,p6,p7, p8)
```


```{r}
#png(height=400, width=800, filename="ggplots3", type="cairo")
plot_grid(p9, p10, p11,p12)
```


```{r}
#png(height=400, width=800, filename="ggplots4", type="cairo")
plot_grid(p13,p14)
```



## Run various correlations based on plots 

```{r}
cor(HighGHI_new$Political_Stability, HighGHI_new$GDP)
```


```{r}
cor(HighGHI_new$Political_Stability, HighGHI_new$PCT_Arable_Land)
```


```{r}
cor(MidGHI_new$Political_Stability, MidGHI_new$PCT_Arable_Land)
```



```{r}
cor(HighGHI_new$Political_Stability, HighGHI_new$PCT_Anemia_Women)
```



## Multicollinearity Checks based on plot below

```{r}
cor(HighGHI_new$Cereal_Import_Dependency, HighGHI_new$AVG_Food_Production)
```


```{r}
cor(HighGHI_new$Undernourishment, HighGHI_new$AVG_Protein_Supply)
```

```{r}
cor(HighGHI_new$PC_Food_Production_Variability, HighGHI_new$AVG_Food_Production)
```


```{r}
cor(HighGHI_new$Undernourishment, HighGHI_new$AVG_Dietary_Esupply)
```


```{r}
cor(HighGHI_new$Cereal_Import_Dependency, HighGHI_new$PCT_Obesity_18Older)
```


## Correlation Plot

```{r}

library(corrplot)


HighGHI_cor <- cor(HighGHI_new)

corrplot(HighGHI_cor, tl.col = "black")
```


## Model with all variables 

```{r}
H_GHI_model= lm(Political_Stability ~ AVG_Dietary_Esupply + AVG_Food_Production + AVG_Protein_Supply + AVG_Protein_Supply_AO + GDP + Undernourishment + Cereal_Import_Dependency + PCT_Arable_Land + Imports_vs_Exports + PC_Food_Production_Variability + PC_Food_Supply_Variability + PCT_Basic_Drinking_Water + PCT_Obesity_18Older + PCT_Anemia_Women, data= HighGHI_new )
```


```{r}
summary(H_GHI_model)
```


## Best Subset Selection

```{r}
library(leaps)
leaps <- regsubsets(Political_Stability ~ AVG_Dietary_Esupply + AVG_Food_Production + 
    AVG_Protein_Supply + AVG_Protein_Supply_AO + GDP + Undernourishment + 
    Cereal_Import_Dependency + PCT_Arable_Land + Imports_vs_Exports + 
    PC_Food_Production_Variability + PC_Food_Supply_Variability + 
    PCT_Basic_Drinking_Water + PCT_Obesity_18Older + PCT_Anemia_Women, 
    data = HighGHI_new)

BestAdjR2 = which.max(summary(leaps)$adjr2)

coef(leaps,BestAdjR2)
```

## Final model

```{r}
library(ggplot2)
library(sjPlot)


H_GHI_Final <- lm(Political_Stability ~  AVG_Food_Production + AVG_Protein_Supply_AO + GDP +  PCT_Arable_Land  + Imports_vs_Exports +  PCT_Obesity_18Older+ PCT_Anemia_Women, data= HighGHI_new )


summary(H_GHI_Final)

tab_model(H_GHI_Final)


```



#Refined model:
$\widehat{Political Stability} = -.3300 - .001811*AVGFoodProduction - .01860*AVGProteinSupplyAnimalOrigin + .0001013*GDP - .01670PCTArableLand - .001294ImportsVSExports + .03010*PCTObesity18orOlder -.01998*PCTAnemiaWomen$


```{r}
#MSE <-mean(H_GHI_Final$residuals^2)
#print(MSE)
```




## Anova between bigger and smaller model

```{r}

anova(H_GHI_Final)
anova(H_GHI_model)


anova(H_GHI_Final, H_GHI_model)

## Adding variables did not lead to a better model fit
## p>.05 choose reduced model; fail to reject the null which is the smaller model
```




## Moderate to low GHI model with same variables as final model for High GHI


```{r}
M_GHI_Final <- lm(Political_Stability ~   AVG_Food_Production + AVG_Protein_Supply_AO + GDP +  PCT_Arable_Land  + Imports_vs_Exports +  PCT_Obesity_18Older+ PCT_Anemia_Women, data= MidGHI_new )
summary(M_GHI_Final)

## Lower R2 than High GHI model
```



## Check for Normality of High GHI model


```{r}
library(olsrr)

ols_test_normality(H_GHI_Final)
ols_test_correlation(H_GHI_Final)
ols_plot_resid_fit(H_GHI_Final)

```

```{r}
pdf("Histogram.pdf", width = 7, height =5)
m <- ols_plot_resid_hist(H_GHI_Final)
m + theme_grey(base_size = 14)
```


```{r}
pdf("QQ.pdf", width = 7, height =5)
n <- ols_plot_resid_qq(H_GHI_Final)
n + theme_grey(base_size = 14)
```



## Load in combined dataset


```{r}
library(readxl)

GHI <- read_excel("C:/Users/mtbri/OneDrive - Nexus365/Summatives/GHI_combined_12_27.xlsx")
```


## Summary of variables and omit missing values 

```{r}
GHI_new = GHI
summary(GHI)
GHI_new <- na.omit(GHI_new) 

```


## Plot all countries with varying degrees of undernourishment against levels of political stability

```{r}
pdf("rplot.pdf", width = 11, height =8)
```


```{r}
library(ggrepel)
l <-ggplot(GHI_new, aes(x = Undernourishment, y = Political_Stability, color = High, label = Country) ) + labs(x= "Undernourishment", y= "Political Stability") +
     geom_point() + geom_text_repel()
l + theme_grey(base_size = 14)
dev.off()
```


## Remove country to create linear regression models based on continuous variables

```{r}
GHI_new$Country <- NULL
```




## Significant Variable Comparisons with Confidence Intervals


```{r}
Anemia_model_low <- lm(Political_Stability ~  PCT_Anemia_Women, data= MidGHI_new)

confint(Anemia_model_low)


Anemia_model_high <- lm(Political_Stability ~  PCT_Anemia_Women, data= HighGHI_new)

confint(Anemia_model_high)

GDP_model_low <- lm(Political_Stability ~  GDP, data= MidGHI_new)

confint(GDP_model_low)


GDP_model_high <- lm(Political_Stability ~  GDP, data= HighGHI_new)

confint(GDP_model_high)



Arable_land_model_low <- lm(Political_Stability ~  PCT_Arable_Land, data= MidGHI_new)

confint(Arable_land_model_low)


Arable_land_model_high <- lm(Political_Stability ~ PCT_Arable_Land, data= HighGHI_new)

confint(Arable_land_model_high)


```

## Additional Plots to Visualize

```{r}
#pdf("Combined_1.pdf", width = 7, height =5)
library(ggplot2)

fitlm = lm(Political_Stability ~  PCT_Anemia_Women, data= MidGHI_new)
fitlm2 = lm(Political_Stability ~  PCT_Anemia_Women, data= HighGHI_new)

g <- ggplot(GHI_new, aes(x = PCT_Anemia_Women, y = Political_Stability, color = High) ) + labs(y= "Political Stability", x= "Prevalence of Anemia Among Women of Reproductive Age")+
     geom_point() + geom_abline(slope = fitlm$coefficients[2], intercept = fitlm$coefficients[1], col = "black", size = 1.5) + geom_abline(slope = fitlm2$coefficients[2], intercept = fitlm2$coefficients[1], col = "blue", size =1.5) + geom_smooth(method= lm, se = FALSE, col = "magenta", size = 1.5)
g + theme_grey(base_size = 14)
#dev.off()
  

  
```


```{r}
#pdf("Combined_2.pdf", width = 7, height =5)
fitlm = lm(Political_Stability ~  PCT_Arable_Land, data= MidGHI_new)
fitlm2 = lm(Political_Stability ~  PCT_Arable_Land, data= HighGHI_new)

h <-ggplot(GHI_new, aes(x = PCT_Arable_Land, y = Political_Stability, color = High) ) +labs(y= "Political Stability", x= "Percent of Arable Land Equipped for Irrigation")+
     geom_point() + geom_abline(slope = fitlm$coefficients[2], intercept = fitlm$coefficients[1], col = "black", size = 1.5) + geom_abline(slope = fitlm2$coefficients[2], intercept = fitlm2$coefficients[1], col = "blue", size =1.5) + geom_smooth(method= lm, se = FALSE, col = "magenta", size = 1.5)
  
  
h + theme_grey(base_size = 14)

#dev.off()
```

```{r}
slope = fitlm$coefficients[2]
slope2 = fitlm2$coefficients[2]

slope-slope2
```



```{r}
#pdf("Combined_3.pdf", width = 7, height =5)
fitlm = lm(Political_Stability ~  GDP, data= MidGHI_new)
fitlm2 = lm(Political_Stability ~  GDP, data= HighGHI_new)

j <- ggplot(GHI_new, aes(x = GDP, y = Political_Stability, color = High)) + labs(y= "Political Stability", x= "GDP")+
     geom_point() + geom_abline(slope = fitlm$coefficients[2], intercept = fitlm$coefficients[1], col = "black", size = 1.5) + geom_abline(slope = fitlm2$coefficients[2], intercept = fitlm2$coefficients[1], col = "blue", size =1.5) + geom_smooth(method= lm, se = FALSE, col = "magenta", size = 1.5)
  
j + theme_grey(base_size = 14) 
  
#dev.off()
```

